### 2025-06-07 lecturenotes_aima104/ch2_aima104.tex
linear_transform

\section{Matrix of a Linear Transformation}
\begin{theorem}
  Let $A$ be an $m \times n$ matrix. The mapping $x \mapsto Ax$ is linear from $\mathbb{R}^n$ to $\mathbb{R}^m$.\\
  Conversely, for a linear map $T: \mathbb{R}^n \to \mathbb{R}^m$ there exists an $m \times n$ matrix $A$ such that $T(\vec{x})=A\vec{x}$ for all $\vec{x} \in \mathbb{R}^n$. In fact, $A$ is the $m \times n$ matrix whose $jth$ column is the vector $T(e_j)$, where $\{e_j\, : \, j=1,2,...,n\}$ is the the basis of the domain $\mathbb{R} ^n$,i.e $A= \left [ T(e_1)\; ... \; T(e_n)\right ]$.
\end{theorem}
\begin{example}
  Describe the linear mapping that has the matrix $\begin{bmatrix}
    0 & -1 \\ 1 & 0
  \end{bmatrix}$. \\[1mm]
  Let the given matrix is denoted by $A$. $A$ is a $2 \times 2$ So, the map is from $\mathbb{R}^2 \mapsto \mathbb{R}^2$. Then for every $\vec{x}=(x_1,x_2) \in \mathbb{R}^2$, Now $A\vec{x}= (-x_2,x_1)$ which is a \textbf{rotation} transformation that rotates every vector through the angle $90^0$ counterclockwise.
\end{example}

\subsection{Exercise}
Find the standard matrix of the linear transformation $T$.
\begin{enumerate}
\item $T(x_1,x_2) = (2x_2-3x_1, x_1-4x_2, 0, x_2) $
\item $T:\mathbb{R} ^2 \mapsto \mathbb{R} ^2$ first reflects points through the $x_1$-axis then reflects points through the line $x_2 = x_1$
\item $T:\mathbb{R} ^2 \mapsto \mathbb{R} ^2$ that rotates each point through an angle $\phi$, with counterclockwise rotation.

\item Describe the transformation of the following matrices geometrically:
  $$\begin{bmatrix}
    1 & 0\\
    0 & -1
  \end{bmatrix}, \hspace{5mm} \begin{bmatrix}
    0 & 1 \\
    1 & 0
  \end{bmatrix}, \hspace{5mm} \begin{bmatrix}
    2 & 0 \\
    0 & 2
  \end{bmatrix}, \hspace{5mm} \begin{bmatrix}
    1 & 0 \\
    0 & 0
  \end{bmatrix}, \hspace{5mm} \begin{bmatrix}
    1 & 0 \\
    1 & 1
  \end{bmatrix}, \hspace{5mm} \begin{bmatrix}
    1 & 3 \\
    0 & 1
  \end{bmatrix}$$
\end{enumerate}

@@

\item The linear transformation $T: V \to W$ is onto if, $dim(R(T) = dim(W)$.

\end{enumerate}

>>>>

\section{Matrix of a Linear Transformation}
\begin{theorem}
  Let $A$ be an $m \times n$ matrix. The mapping $x \mapsto Ax$ is linear from $\mathbb{R}^n$ to $\mathbb{R}^m$.\\
  Conversely, for a linear map $T: \mathbb{R}^n \to \mathbb{R}^m$ there exists an $m \times n$ matrix $A$ such that $T(\vec{x})=A\vec{x}$ for all $\vec{x} \in \mathbb{R}^n$. In fact, $A$ is the $m \times n$ matrix whose $jth$ column is the vector $T(e_j)$, where $\{e_j\, : \, j=1,2,...,n\}$ is the the basis of the domain $\mathbb{R} ^n$,i.e $A= \left [ T(e_1)\; ... \; T(e_n)\right ]$.
\end{theorem}

\subsection{Exercise}
Find the standard matrix of the linear transformation $T$.
\begin{enumerate}
\item For the dilation transformation $T(x)=3x$, $x \in \mathbb{R}^2$.

\item Show that the transformation that rotates each point in $\mathbb{R}^2$, about the origin, through an angle $\phi$ counterclockwise, is a linear transformation by find the standard matrix of the transformation.

\item $T(x_1,x_2) = (2x_2-3x_1, x_1-4x_2, 0, x_2) $
\item $T:\mathbb{R} ^2 \mapsto \mathbb{R} ^2$ first reflects points through the $x_1$-axis then reflects points through the line $x_2 = x_1$
\item $T:\mathbb{R} ^2 \mapsto \mathbb{R} ^2$ that rotates each point through an angle $\phi$, with counterclockwise rotation.

\item Describe the transformation of the following matrices geometrically:
  $$\begin{bmatrix}
    1 & 0\\
    0 & -1
  \end{bmatrix}, \hspace{5mm} \begin{bmatrix}
    0 & 1 \\
    1 & 0
  \end{bmatrix}, \hspace{5mm} \begin{bmatrix}
    2 & 0 \\
    0 & 2
  \end{bmatrix}, \hspace{5mm} \begin{bmatrix}
    1 & 0 \\
    0 & 0
  \end{bmatrix}, \hspace{5mm} \begin{bmatrix}
    1 & 0 \\
    1 & 1
  \end{bmatrix}, \hspace{5mm} \begin{bmatrix}
    1 & 3 \\
    0 & 1
  \end{bmatrix}$$
\end{enumerate}
\begin{example}
  Describe the linear mapping that has the matrix $\begin{bmatrix}
    0 & -1 \\ 1 & 0
  \end{bmatrix}$. \\[1mm]
  Let the given matrix is denoted by $A$. $A$ is a $2 \times 2$ So, the map is from $\mathbb{R}^2 \mapsto \mathbb{R}^2$. Then for every $\vec{x}=(x_1,x_2) \in \mathbb{R}^2$, Now $A\vec{x}= (-x_2,x_1)$ which is a \textbf{rotation} transformation that rotates every vector through the angle $90^0$ counterclockwise.
\end{example}

\subsection{Types of Transformation}
Geometrically, linear transformation are basically of the following types:
\begin{enumerate}
\item Reflection
\item Rotation
\item Contraction and Expansion
\item Shears: Horizontal and Vertical
\item Projections
\end{enumerate}

@@

\item The linear transformation $T: V \to W$ is onto if, $dim(R(T) = dim(W)$.
\end{enumerate}

\section{Properties of Linear Transformation}
Let $T: \mathbb{R}^n \to \mathbb{R}^m$ be a linear transformation.
\begin{enumerate}
\item Then $T(0)=0$.
\item $T$ is one-to-one if and only if the equation $T(x)=0$ has only the trivial solution. (\textit{Indirectly it has got to do with the nullity or nullspace of the matrix.})
\item $T$ is onto if and only if the columns of the corresponding standard matrix spans the $\mathbb{R}^m$. (\textit{So it has got to do with the column space of the matrix.})
\end{enumerate}
Give an example of one-to-one and not one-to-one linear transformation.
